Hierarchical Large Language Models for Multi-task problems

### 2024-03-04 paper/main.tex
more writing on lora++

\usepackage[
backend=biber,
style=alphabetic,
sorting=ynt
]{biblatex}
\usepackage{amsmath}
\usepackage{booktabs}
\addbibresource{references.bib}

@@

@@

\subsection{Pretrained Autoregressive Language Model}
Our foundational model is a pretrained autoregressive language model, denoted by its parameters $\theta_{full}$ and its architecture, referred to as $LLM$. The model's probability distribution is given by:

\begin{equation} \label{eq:LLMprob}
P(w_{1:W} | \theta_{full}) = \prod_{i=1}^{W-1} LLM(w_{i+1} | w_{1:i})
\end{equation}
%
Here, $\theta_{full}$ represents the set of parameters obtained from pretraining on a large and diverse dataset, distinct from our current dataset $\mathcal{D}$.
We omit to write $\theta_{full}$ explicitly in the model, as the autoregressive language model always depend on these parameters.

\subsection{The Hierarchical LoRA model}
For each task $d$ we have a LORA parameter set $\theta_d$ that is used to fine-tune to that task.
Hence, the likelihood of our problem follows from combining equation \ref{eq:data} and \ref{eq:LLMprob}:
\begin{equation} \label{eq:lik}
    \text{L}(\mathcal{D} | \theta_1, \theta_2, \ldots, \theta_D) = \prod_{d=1}^D L(\mathcal{D}_d | \theta_d) = \prod_{d=1}^D \prod_{n=1}^{N_d} \prod_{i=1}^{W_n-1} \text{LLM}(w_{d,n,i+1} | w_{d,n,1:i}; \theta_d)
\end{equation}
%
%
Each task will share some information with each other, and we model this by introducing a Gaussian hierarchical prior over the task parameters $\theta_{1:D}$, given by
\begin{equation} \label{eq:prior}
    P(\theta_{1:D} | \Theta, \tau ) = \prod_{d=1}^D N(\theta_d ; \Theta, \frac{1}{\tau} I)
\end{equation}
%
where $\Theta$ is a set of hierarchical mean parameters, $I$ is a unit diagonal matrix and $\tau \geq 0$ is a scalar precision hyperparameter controlling how similar task parameters $\theta_d$ should be to the hierarchical mean parameters $\Theta$.
We let $\Theta$ have a uniform prior (i.e. $P(\Theta) = 1$).

>>>>

\usepackage[
backend=biber,
style=alphabetic,
sorting=ynt
]{biblatex}
\usepackage{amsmath}
\usepackage{amssymb}
\newcommand{\R}{\mathbb{R}}
\usepackage{booktabs}
\addbibresource{references.bib}

@@

\subsection{Multi-task Learning}
Multi-task Learning \cite{caruana_multitask_nodate} is a popular machine learning approach for training one model on multiple tasks. A common approach is to share the lower layers of a neural network between tasks, and have separate output layers for each task. Our paper focuses on a different approach, where we share information between tasks through a hierarchical prior over the task-specific parameters.

@@

\subsection{Pretrained Autoregressive Language Model}
Our foundational model is a pretrained autoregressive language model, denoted by its parameters $\theta_{full}$ and its architecture, referred to as a Large Language Model ($LLM$). The parameters $\theta_{full}$ has been obtained from pretraining on a large and diverse dataset, distinct from our current dataset $\mathcal{D}$. The model's probability distribution is given by:
The model's probability distribution is given by:

\begin{equation} \label{eq:LLMprob}
P(w_{1:W} | \theta_{full}) = \prod_{i=1}^{W-1} LLM(w_{i+1} | w_{1:i})
\end{equation}
%
where $LLM(w_{i+1} | w_{1:i})$ is the probability of the next token $w_{i+1}$ given the previous tokens $w_{1:i}$ in the document.
We omit to write $\theta_{full}$ explicitly in the model, as the autoregressive language model always depend on these parameters.

\subsection{Low-Rank Adaption (LoRA)}
A Large Language Model contains many linear layers, and the LoRA method reduces the number of trainable parameters by reparameterizing the weight parameters ($W \in \R^{n_1 \times n_2}$) of these layers into the original pretrained weights $W_{full}$ and a low-rank decomposition $A$ and $B$ (\cite{hayou_lora_2024}):
\begin{equation} \label{eq:LoRA}
    W = W_{full} + \frac{\alpha}{r} BA 
\end{equation}
%
where $W_{full} \in \R^{n_1 \times n_2}$ is the original pretrained weight, $B \in \R^{n_1 \times r}$ and $A \in \R^{r \times n_2}$ are the low-rank factors, $r \ll \text{min}(n_1, n_2)$ is the low rank dimension and $\alpha$ is a scaling hyperparameter.

During LoRA fine-tuning, the low-rank factors $A$ and $B$ are learned as normal parameters, while the original pretrained weights $W_{full}$ are kept fixed.

\subsection{The Hierarchical LoRA model}
In the hierarchical LoRA model, we introduce a LoRA parameter set for each task $d$, denoted as $\theta_d$, containing all the low-rank decomposition matricies $A$ and $B$ from each decomposed layer in the Large Language Model. The likelihood of the data $\mathcal{D}$ given the task parameters $\theta_{1:D}$ is then given by (combining equation \ref{eq:data} and \ref{eq:LLMprob}):
\begin{equation} \label{eq:lik}
    \text{L}(\mathcal{D} | \theta_1, \theta_2, \ldots, \theta_D) = \prod_{d=1}^D L(\mathcal{D}_d | \theta_d) = \prod_{d=1}^D \prod_{n=1}^{N_d} \prod_{i=1}^{W_n-1} \text{LLM}(w_{d,n,i+1} | w_{d,n,1:i}; \theta_d)
\end{equation}
%
%
Each task will share some information with each other, and we model this by introducing a Gaussian hierarchical prior over the task parameters $\theta_{1:D}$, given by
\begin{equation} \label{eq:prior}
    P(\theta_{1:D} | \Theta, \tau ) = \prod_{d=1}^D N(\theta_d ; \Theta, \frac{1}{\tau} I)
\end{equation}
%
where $\Theta$ is a set of hierarchical mean parameters, $I$ is a unit diagonal matrix and $\tau \geq 0$ is a scalar precision hyperparameter controlling how similar task parameters $\theta_d$ should be to the hierarchical mean parameters $\Theta$.
We let $\Theta$ have a uniform prior (i.e. $P(\Theta) = 1$).
